---
title: "Synchrony investigations - Fagus masting"
author: "Victor, Lizzie"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(terra) # process raster data
library(dplyr)
library(ggplot2)
library(ggforce) # geom_link2 function
library(scales) # date on x-axis
library(mgcv) # gam

wd <- "C:/Users/vandermeersch/Documents/CEFE/projects/synchrony"

# Data from Journe et al. 2024 ()
## note that this is the filtered dataset used by V. Journe, we may want to start from scratch with raw MasTree data
journe_d <- read.csv(file.path(wd, "data/journe2024", "mastree_subset_fagus.csv")) 

```

### Get E-OBS climate data 

```{r eobs}

rerun <- FALSE # switch because computation is too long

if(rerun){
  eobs_r <- rast(file.path(wd, "data/eobs", "tg_ens_mean_0.1deg_reg_v29.0e.nc"))
  
  # Loop on sites to get E-OBS daily temperature for the years recorded
  process_data <- data.frame()
  for(s in unique(journe_d$sitenewname)){
    
    lat <- unique(journe_d[journe_d$sitenewname == s, "Latitude"])
    lon <- unique(journe_d[journe_d$sitenewname == s, "Longitude"])
    years <- journe_d[journe_d$sitenewname == s, "Year"]
    years <- c(min(years)-2, min(years-1), years) # two years before seed fall
    
    sub_r <- subset(eobs_r, which(time(eobs_r, format = "years") %in% years))
    tmean <- terra::extract(
      sub_r, 
      matrix(c(lon, lat), nrow = 1))
    
    process_data <- rbind(process_data, data.frame(Latitude = lat, Longitude = lon, sitenewname = s, Tmean = t(tmean), 
                                                   Date = time(sub_r), Year = time(sub_r, format = "years")))
  }
  
  # Add annual seed production (including one and two years later) -- not a very elegant way
  journe_T2 <- journe_d[,c("Longitude", "Latitude", "sitenewname", "Year", "Value")]
  journe_T2$Year <- journe_T2$Year-2
  names(journe_T2)[5] <- "Value2"
  journe_T1 <- journe_d[,c("Longitude", "Latitude", "sitenewname", "Year", "Value")]
  journe_T1$Year <- journe_T1$Year-1
  names(journe_T1)[5] <- "Value1"
  process_data <- process_data %>%
    dplyr::left_join(journe_d[c("Longitude", "Latitude", "sitenewname", "Year", "Value")], 
                     by = c("Longitude", "Latitude", "sitenewname", "Year")) %>%
    dplyr::left_join(journe_T1[c("Longitude", "Latitude", "sitenewname", "Year", "Value1")], 
                     by = c("Longitude", "Latitude", "sitenewname", "Year")) %>% 
    dplyr::left_join(journe_T2[c("Longitude", "Latitude", "sitenewname", "Year", "Value2")], 
                     by = c("Longitude", "Latitude", "sitenewname", "Year"))
  
  saveRDS(process_data, file = file.path(wd, "data/processed", "journe2024_replicate.rds"))
  
}else{
  
  process_data <- readRDS(file.path(wd, "data/processed", "journe2024_replicate.rds"))
  
}


```

### Computing the "moving window correlation"

From my understanding, the correlation is computed as follow.

```{r journe_corr}

## from Journe et al. "a moving Spearman correlation between 
## log-transformed annual seed production and mean daily temperature"

dat <- data.frame(
  site = process_data$sitenewname,
  doy = as.numeric(strftime(process_data$Date, format = "%j")),
  tmean = process_data$Tmean,
  prod0 = process_data$Value,
  prod1 = process_data$Value1,
  prod2 = process_data$Value2
)

# Moving window correlation across year for each site (note the log-transformation of seed production)
mwcor <- lapply(7:365, function(d) {

  dat %>%
    dplyr::filter(doy >= (d-6) & doy <= d) %>% # 7-day window
    dplyr::select(-doy) %>%
    group_by(site) %>%
    reframe(
      cor0 = cor(tmean, log(prod0), use="complete.obs", method = "spearman"),
      cor1 = cor(tmean, log(prod1), use="complete.obs", method = "spearman"),
      cor2 = cor(tmean, log(prod2), use="complete.obs", method = "spearman"),
      doy = d) %>%
    as.data.frame()

})

mwcor <- as.data.frame(do.call(rbind, mwcor)) %>%
  tidyr::pivot_longer(cols = c("cor0", "cor1", "cor2")) %>% 
  # "trick" to plot in the correct order
  dplyr::mutate(doy_plot = if_else(name == "cor2", as.Date(doy, origin = "2001-01-01"),
                                   as.Date(ifelse(name == "cor1", as.Date(doy, origin = "2002-01-01"), 
                                                  as.Date(doy, origin = "2003-01-01")))))

```

```{r fig1c}

mwcor %>%
  group_by(doy_plot) %>%
  summarise(cor = mean(value)) %>%
  ggplot(aes(x=doy_plot)) + 
  annotate("rect", xmin=as.Date("2001-01-01"), xmax=as.Date("2002-01-01"), 
           ymin=-Inf, ymax=Inf, alpha=0.2, fill="lightgrey") +
  annotate("rect", xmin=as.Date("2003-01-01"), xmax=as.Date("2003-11-01"), 
           ymin=-Inf, ymax=Inf, alpha=0.2, fill="lightgrey") +
  ggforce::geom_link2(lwd=0.3, n = 50,
             aes(y = cor, colour = after_stat(y < 0)))+
  scale_colour_manual(values = c("darkblue", "darkred")) + 
  ggh4x::stat_difference(aes(ymin = 0, ymax = cor), alpha = 0.3) +
  scale_fill_manual(values = c("darkblue", "darkred")) + 
  scale_x_date(date_labels = "%d-%m",
               breaks = as.Date(c("2001-06-01", "2002-01-01", "2002-06-01", "2003-01-01", "2003-06-01")),
               limits = as.Date(c(NA, "2003-11-01")), expand = c(0,0)) +
  scale_y_continuous(limits = c(-0.35, 0.35), expand = c(0,0)) +
  geom_vline(data = data.frame(sol = as.Date(c("2001-06-21", "2002-06-21"))), 
                               aes(xintercept = sol), linetype = "dashed") +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Spearman corr.", x = "Time")

```

But we could also compute these correlations differently, right?
If we take the correlation between moving average temperature and seed production, it gives quite similar results.

```{r fig1c_alternative}

mwcor_alt <- dat %>%
  group_by(site) %>%
  dplyr::mutate(rwtmean = zoo::rollapplyr(tmean, 7 , mean, fill = NA)) %>% 
  ungroup() %>%
  group_by(site, doy) %>%
  dplyr::summarise(
    cor0 = cor(rwtmean, prod0, use="complete.obs", method = "spearman"),
    cor1 = cor(rwtmean, prod1, use="complete.obs", method = "spearman"),
    cor2 = cor(rwtmean, prod2, use="complete.obs", method = "spearman")
  ) %>%
  tidyr::pivot_longer(cols = c("cor0", "cor1", "cor2")) %>% 
  dplyr::mutate(doy_plot = if_else(name == "cor2", as.Date(doy, origin = "2001-01-01"),
                                   as.Date(ifelse(name == "cor1", as.Date(doy, origin = "2002-01-01"), 
                                                  as.Date(doy, origin = "2003-01-01")))))

mwcor_alt %>%
  group_by(doy_plot) %>%
  summarise(cor = mean(value)) %>%
  ggplot(aes(x=doy_plot)) +  
  annotate("rect", xmin=as.Date("2001-01-01"), xmax=as.Date("2002-01-01"), 
           ymin=-Inf, ymax=Inf, alpha=0.2, fill="lightgrey") +
  annotate("rect", xmin=as.Date("2003-01-01"), xmax=as.Date("2003-11-01"), 
           ymin=-Inf, ymax=Inf, alpha=0.2, fill="lightgrey") +
  ggforce::geom_link2(lwd=0.3, n = 50,
             aes(y = cor, colour = after_stat(y < 0)))+
  scale_colour_manual(values = c("darkblue", "darkred")) + 
  ggh4x::stat_difference(aes(ymin = 0, ymax = cor), alpha = 0.3) +
  scale_fill_manual(values = c("darkblue", "darkred")) + 
  scale_x_date(date_labels = "%d-%m",
               breaks = as.Date(c("2001-06-01", "2002-01-01", "2002-06-01", "2003-01-01", "2003-06-01")),
               limits = as.Date(c(NA, "2003-11-01")), expand = c(0.01,0.01)) +
  scale_y_continuous(limits = c(-0.35, 0.35), expand = c(0,0)) +
  geom_vline(data = data.frame(sol = as.Date(c("2001-06-21", "2002-06-21"))), 
                               aes(xintercept = sol), linetype = "dashed") +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Spearman corr. (alternative)", x = "Time")

```

### Looking at day length effect

```{r daylength}

# Computing day-length

## we need latitude
mwcor <- mwcor %>%
  dplyr::left_join(unique(journe_d[c("Latitude", "sitenewname")]), 
                   by = c("site" = "sitenewname")) %>%
  dplyr::mutate(lat = Latitude) %>% dplyr::select(-Latitude)

doy <- mwcor$doy
phi <- (mwcor$lat*pi)/180 # in radian

## solar declination
delta <- 0.409 * sin(2 * pi/365 * doy - 1.39)

## solar hour angle
aux <- -tan(phi) * tan(delta) 
# (handle particular cases)
aux[aux > 1] <- 1 
aux[aux < -1] <- -1
wo <- acos(-tan(phi)*tan(delta))

## daylength
dl <- 24/pi*wo
mwcor$dl <- dl

mwcor <- mwcor %>%
  dplyr::group_by(site) %>%
  dplyr::mutate(dlnorm = (dl-min(dl))/(max(dl)-min(dl))) # "for each site, day length was normalized to the 0–1 range"

```

```{r fig2a}

mwcor %>%
  dplyr::filter(name == "cor1") %>% 
  dplyr::mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)) %>%
  ggplot(aes(x=dlnorm, y = value, color = doy < 172)) + 
  # geom_smooth(method = 'loess') + 
  geom_point(alpha = 0.3, shape=16, size = 1, stroke = 0) + 
  geom_vline(data = data.frame(sol = 0), 
             aes(xintercept = sol), linetype = "dashed") +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Spearman corr.", x = "Day length (scaled by site)") +
  scale_x_continuous(breaks = c(-1,-0.5,0,0.5,1), labels = c(0,0.5,1,0.5,0))

```

With the "alternative" way of computing correlation (moving window temperature):

```{r fig2a_alternative}

# Computing day-length

## we need latitude
mwcor_alt <- mwcor_alt %>%
  dplyr::left_join(unique(journe_d[c("Latitude", "sitenewname")]), 
                   by = c("site" = "sitenewname")) %>%
  dplyr::mutate(lat = Latitude) %>% dplyr::select(-Latitude)

doy <- mwcor_alt$doy
phi <- (mwcor_alt$lat*pi)/180 # in radian

## solar declination
delta <- 0.409 * sin(2 * pi/365 * doy - 1.39)

## solar hour angle
aux <- -tan(phi) * tan(delta) 
# (handle particular cases)
aux[aux > 1] <- 1 
aux[aux < -1] <- -1
wo <- acos(-tan(phi)*tan(delta))

## daylength
dl <- 24/pi*wo
mwcor_alt$dl <- dl

mwcor_alt <- mwcor_alt %>%
  dplyr::group_by(site) %>%
  dplyr::mutate(dlnorm = (dl-min(dl))/(max(dl)-min(dl))) # "for each site, day length was normalized to the 0–1 range"

mwcor_alt$dl <- dl

mwcor_alt <- mwcor_alt %>%
  dplyr::group_by(site) %>%
  dplyr::mutate(dlnorm = (dl-min(dl))/(max(dl)-min(dl))) # "for each site, day length was normalized to the 0–1 range"

mwcor_alt %>%
  dplyr::filter(name == "cor1") %>% 
  dplyr::mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)) %>%
  ggplot(aes(x=dlnorm, y = value, color = doy < 172)) + 
  # geom_smooth(method = 'loess') + 
  geom_point(alpha = 0.3, shape=16, size = 1, stroke = 0) + 
  geom_vline(data = data.frame(sol = 0), 
             aes(xintercept = sol), linetype = "dashed") +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Spearman corr.", x = "Day length (scaled by site)") +
  scale_x_continuous(breaks = c(-1,-0.5,0,0.5,1), labels = c(0,0.5,1,0.5,0))

```

### Looking at GAM-related stuff

See if we can fit a similar GAM than Journe et al.

```{r fig2a_gam}

data_fit <- mwcor_alt %>%
  dplyr::filter(name == "cor1") %>% 
  dplyr::mutate(
    source = strsplit(site,split='_')[[1]][1],
    prepost = factor(doy < 172)) %>%
  # "we rescaled the Spearman correlations to the 0–1 range. The scaling was done for each site separately" 
  dplyr::group_by(site) %>%
  dplyr::mutate(
    maxv = max(value), minv = min(value),
    valuenorm = (value-minv)/(maxv-minv))

# dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)

journe_gam <- mgcv::gam(
  formula = valuenorm ~ prepost + s(dlnorm, by = prepost),
  data = data_fit,
  family=betar(link="logit", eps=.Machine$double.eps*1e6),
  method="REML")

pred_gam <- data.frame(
  dlnorm = data_fit$dlnorm,
  doy = data_fit$doy,
  maxv = data_fit$maxv,
  minv = data_fit$minv,
  valuenorm = data_fit$valuenorm)%>%
  mutate(pred = predict(journe_gam, type="response",se.fit = FALSE)* (maxv-minv) + minv) %>%
  group_by(doy) %>%
  summarise(pred = mean(pred), dlnorm = mean(dlnorm))
  
ggplot() +
  geom_point(
    data = data_fit %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = dlnorm, y = value), size = 0.1, alpha = 0.1) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = dlnorm, y = pred), color = "white", linewidth = 0.7) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = dlnorm, y = pred), color = "#0B3954", linewidth = 0.7) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  labs(y = "Spearman corr.", x = "Day length (scaled by site)") +
  scale_x_continuous(breaks = c(-1,-0.5,0,0.5,1), labels = c(0,0.5,1,0.5,0)) +
  coord_cartesian(expand = FALSE)

```

If we plot according to DOY (to avoid squeezing the data around the solstice...), the big step we assume at solstice seems much more bizarre.

```{r fig2a_gam_doy}

# pred_gam <- data.frame(
#   dlnorm = data_fit$dlnorm,
#   doy = data_fit$doy,
#   maxv = data_fit$maxv,
#   minv = data_fit$minv,
#   valuenorm = data_fit$valuenorm)%>%
#   mutate(pred = predict(journe_gam, type="response",se.fit = FALSE)* (maxv-minv) + minv) %>%
#   group_by(doy) %>%
#   summarise(pred = mean(pred), dlnorm = mean(dlnorm))
  
ggplot() +
  geom_point(
    data = data_fit %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = value), size = 0.1, alpha = 0.1) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.7) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#0B3954", linewidth = 0.7) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  labs(y = "Spearman corr.", x = "DOY") +
  coord_cartesian(expand = FALSE)

```

And if we look at the residuals/time... We find this wavy pattern.

```{r fig2a_gam_res}

data.frame(
  doy = data_fit$doy,
  res = residuals(journe_gam)
) %>%
  ggplot(aes(x = doy)) +
  geom_point(aes(y = res), size = 0.1, alpha = 0.1) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Residuals", x = "DOY")

```

We may want to try to fit a GAM with AR1 correlation? Still assuming a priori a solstice effect as Journe et al.
Still plotting against DOY to avoid squeezing stuff around summer solstice.

```{r fig2a_gam_wTC}

journe_gam_AR1 <- mgcv::gamm(
  formula = value ~ prepost + s(dlnorm, by = prepost),
  correlation = corAR1(form = ~ doy | site),
  data = data_fit,
  method="REML")

pred_gam_AR1 <- data.frame(
  dlnorm = data_fit$dlnorm,
  doy = data_fit$doy,
  maxv = data_fit$maxv,
  minv = data_fit$minv,
  value = data_fit$value)%>%
  mutate(pred = predict(journe_gam_AR1$gam, type="response",se.fit = FALSE)) %>%
  group_by(doy) %>%
  summarise(pred = mean(pred), dlnorm = mean(dlnorm))
  
ggplot() +
  geom_point(
    data = data_fit %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = value), size = 0.1, alpha = 0.1) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.7) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#0B3954", linewidth = 0.7) +
  geom_line(
    data = pred_gam_AR1 %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.9) +
  geom_line(
    data = pred_gam_AR1%>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)) ,
    aes(x = doy, y = pred), color = "#A5151C", linewidth = 0.7) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  labs(y = "Spearman corr.", x = "DOY") +
  coord_cartesian(expand = FALSE)


```

What if we remove the solstice-thingy variable?

```{r fig2a_gam_wTC_woSol}

# data_fit$prepost <- factor(data_fit$doy < 195)

basic_gam_AR1 <- mgcv::gamm(
  formula = value ~ s(dlnorm),
  correlation = corAR1(form = ~ doy | site),
  data = data_fit %>% dplyr::mutate(site = factor(site)),
  method="REML")


pred_gam_AR1 <- data.frame(
  dlnorm = data_fit$dlnorm,
  doy = data_fit$doy,
  maxv = data_fit$maxv,
  minv = data_fit$minv,
  value = data_fit$value)%>%
  mutate(pred = predict(basic_gam_AR1$gam, type="response",se.fit = FALSE)) %>%
  group_by(doy) %>%
  summarise(pred = mean(pred), dlnorm = mean(dlnorm))
  
ggplot() +
  geom_point(
    data = data_fit %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = value), size = 0.1, alpha = 0.1) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.7) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#0B3954", linewidth = 0.7) +
  geom_line(
    data = pred_gam_AR1 %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.9) +
  geom_line(
    data = pred_gam_AR1 %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#A5151C", linewidth = 0.7) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  labs(y = "Spearman corr.", x = "DOY") +
  coord_cartesian(expand = FALSE) +
  geom_vline(data = data.frame(doyp = c(106,228)), 
             aes(xintercept = doyp), linetype = "dotted") 


```

```{r fig2a_gam_res_comp}

basic_gam_noAR <- mgcv::gamm(
  formula = value ~ prepost + s(dlnorm, by = prepost),
  data = data_fit,
  method="REML")


data.frame(
  doy = rep(data_fit$doy,2),
  res = c(resid(basic_gam_noAR$lme, type = "normalized"), resid(basic_gam_AR1$lme, type = "normalized")),
  type = c(rep("noAR - solstice", length(data_fit$doy)), rep("AR1 - no solstice", length(data_fit$doy)))
) %>%
  ggplot(aes(x = doy)) +
  facet_wrap(~ type) + 
  geom_point(aes(y = res), size = 0.2, alpha = 0.2) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank()) +
  labs(y = "Residuals (scaled)", x = "DOY") +
  coord_cartesian(ylim = c(-5,5))

```







```{r fig2a_gam_wTC_woSol}

# data_fit$prepost <- factor(data_fit$doy < 195)

basic_gam_AR1 <- mgcv::gamm(
  formula = value ~ s(dlnorm, by = site),
  correlation = corAR1(form = ~ doy | site),
  data = data_fit %>% dplyr::mutate(site = factor(site)),
  method="REML")


pred_gam_AR1 <- data.frame(
  dlnorm = data_fit$dlnorm,
  doy = data_fit$doy,
  maxv = data_fit$maxv,
  minv = data_fit$minv,
  value = data_fit$value,
  site = data_fit$site) %>%
  mutate(pred = predict(basic_gam_AR1$gam, type="response",se.fit = FALSE)) %>%
  group_by(doy, site) %>%
  summarise(pred = mean(pred), dlnorm = mean(dlnorm))
  
ggplot() +
  geom_point(
    data = data_fit %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = value), size = 0.1, alpha = 0.1) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.7) +
  geom_line(
    data = pred_gam %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#0B3954", linewidth = 0.7) +
  geom_line(
    data = pred_gam_AR1 %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "white", linewidth = 0.9) +
  geom_line(
    data = pred_gam_AR1 %>% mutate(dlnorm = ifelse(doy < 172, -1+dlnorm, 1-dlnorm)),
    aes(x = doy, y = pred), color = "#A5151C", linewidth = 0.7) +
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.minor = element_blank(), panel.grid.major = element_blank()) +
  labs(y = "Spearman corr.", x = "DOY") +
  coord_cartesian(expand = FALSE) +
  geom_vline(data = data.frame(doyp = c(106,228)), 
             aes(xintercept = doyp), linetype = "dotted") 


```